Probing heavy ion collisions using quark and gluon jet substructure

### 2018-02-28 qg_ML.tex
update from raghav, edited conclusions and fig5 update

We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jw Monte Carlo simulations as an example. Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems. We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction. We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets. All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns. Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region. This is closely related to the lost of information in subleading subjets, which is a characteristic feature of the \jw simulation.

The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations. Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications. The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization. In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice to identify robust features of jet-medium interactions. Our work suggests a systematically improvable framework which will serve this important purpose.

>>>>

We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jw Monte Carlo simulations as an example. Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems. We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction. We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets. All the methods studied in this paper perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns. Through multiple methods and observables, we consistently find the dominant feature of the \jw jet quenching model to be the increase of soft particle multiplicity due to medium recoils throughout the jet region. This is closely related to the loss of information in subleading subjets, which is a characteristic feature of \jw.

The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations. Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications. By studying the telescoping deconstruction framework, we suggest new jet substructure observables for jet quenching studies, that we hope will provide qualitatively new insights order-by-order when measured experimentally. In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice in order to identify robust features of jet-medium interactions. Our work represents a thorough systematic framework that we hope will serve this important purpose.
